---
title: "Identifying Fraud Patterns Analysis"
author: "Layla Ashour"
output: html_document
date: "2024-05-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing libraries 

```{r }
library(dplyr)
library(ggplot2)
```
Loading dataset
```{r}
raw_data <- read.csv("Data.csv")
```
Data Exploration 
```{r}
head(raw_data)
summary(raw_data)
str(raw_data)
```
selected data 
```{r}
selected_data <- raw_data %>%
  select("customer", "age", "gender", "zipcodeOri", "category", "amount", "fraud")
```
Data Cleaning & Processing 

no missing data
removed the '' as it was not the correct format 
removed the 0 from the age group as that looks like a mistake or could possibly represent NA values.
removed 0 from the amount because if the amount is 0 its not a valid transaction 
filterns gender to not include values other than F&M, because thats not relavant 

```{r}
missing_values <- sum(is.na(selected_data))

unique_gender <- unique(selected_data$gender)
selected_data$gender <- gsub("'", "",selected_data$gender)
selected_data$age <- gsub("'", "",selected_data$age)
selected_data$category <- gsub("'", "",selected_data$category)
selected_data$zipcodeOri <- gsub("'", "",selected_data$zipcodeOri)
selected_data <- selected_data[selected_data$age != '0', ]
selected_data <- selected_data[selected_data$amount != 0, ]

data <- selected_data %>%
   filter( gender == 'F' | gender == 'M')

```
Statistics  

```{r}
fraud_by_gender <- data %>%
  select(gender,fraud) %>%
  filter(fraud == 1) %>%
  group_by(gender)%>%
  count(gender)

fraud_by_age_group <- data %>%
  select(age,fraud) %>%
  filter(fraud == 1) %>%
  count(age)

```



```{r}
values <- fraud_by_gender$n
labels <- fraud_by_gender$gender

percentages <- round(values / sum(values) * 100) 
labels_percentages <- paste( percentages, "%", sep = "")

label_mapping <- c("M" = "Male", "F" = "Female")
fixed_label <- label_mapping[labels]
  
colors <- c("blue", "green")
pie(values, labels = labels_percentages, col = colors, main = "Fraud by Gender") 

legend("topleft", legend = fixed_label, fill = colors)
```
```{r}

age_group_mapping <- c("1" = "10", "2" = "20","3" = "30","4" = "40","5" = "50","6" = "60")
age_labels <- age_group_mapping
ggplot(fraud_by_age_group, aes(x = age_group_mapping, y = n)) +
geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Fraud by Age Group", x = "Age Group", y = "Count")
```


